### 2016-03-01 conclusions.tex
ready submit

%We also plan to explore other applications of mutant-based metrics.
Many software engineering techniques rely on measuring distance
between program executions.  Not all of these concern failing
executions.  For example, Zhang et al. \cite{issta14} show that FPF based on a simple
Hamming distance over branches covered can significantly improve the
effectiveness of seeded symbolic execution
\cite{Zesti,PersonSeed,BugRedux}.  Mutation response is likely too
expensive to use for this purpose without modification, but sampling
mutants to refine distance when coarser metrics plateau
may be practical.  Mutants may also be useful in determining
execution peers \cite{Sumner2011}, for example to help operators of
spacecraft find similar behaviors to telemetry downloads in testbed
history \cite{KlausRajeev,scriptstospecs}.

>>>>

%We also plan to explore other applications of mutant-based metrics.
Many software engineering techniques rely on measuring distance
between program executions \cite{BallConcept,NearNeighbor}.  Not all
of these concern only failing
executions.  For example, Zhang et al. \cite{issta14} show that FPF based on a simple
Hamming distance over branches covered can significantly improve the
effectiveness of seeded symbolic execution
\cite{Zesti,PersonSeed,BugRedux}.  Mutation response is likely too
expensive to use for this purpose without modification, but sampling
mutants to refine distance when coarser metrics plateau
may be practical.  Mutants may also be useful in determining
execution peers \cite{Sumner2011}, for example to help operators of
spacecraft find similar behaviors to telemetry downlinks in testbed
history \cite{KlausRajeev,scriptstospecs}.

### 2016-03-01 introduction.tex
ready submit

The goal of this paper, an effective metric for computing distance between program executions (in particular, failing test cases), is motivated in large part by progress in automated testing.  Recent advances, some implemented in powerful tools \cite{csmith,jsfunfuzz,ISSTA12,LangFuzz,ZhendongPLDI,ZhendongOOPSLA}, have made it easier to find subtle faults in, e.g., modern optimizing compilers\footnote{In this paper, we take the problem of testing and debugging compilers as representative of the general problem of testing and debugging complex critical systems software.}.  Unfortunately, the blessing of many failing test cases can become a curse for a developer.  Simply discovering which test cases in a large set of failures represent distinct faults (or already known faults) is a very difficult problem \cite{PLDI13,Podgurski04}.  For simple crashes and assertion violations, bucketing failures is sometimes easy.  When a failure is detected by differential testing \cite{Differential}, where two compilers or optimization levels disagree, the problem is so hard that it may turn potential users away from automated test generation \cite{PLDI13}.  Even once a fault is identified, fixing the problem can be very difficult.  Due to complexity of intermediate representation, interaction of optimizations, subtle language semantics, and other issues, debugging compiler faults may often be harder than debugging faults in other kinds of software (this is probably not limited to compilers; debugging systems software in general is hard \cite{mickens}).
We propose the use of program mutants \cite{mutant} to assist debugging based on large sets of automatically generated test cases.  The underlying idea is that if two failures can be repaired in the same way, even if that repair only avoids the failure and does not truly fix it, they are likely due to the same fault. Repairs can also provide  information for understanding and debugging discovered faults. The approach suggests a plausible metric for comparing any two program executions, not just failing test cases.

@@

\subsection{Comparing Executions by Response to Mutation}

@@

The idea of measuring the distance between failures (or other test cases) by their response to mutants is appealing.  The most important aspect of most applications of distances between executions is some aspect of causality \cite{LewisCause,ZellerBook}.  Since at least the work of Hume \cite{Hume1}, causality has been understood as answering some question of the form ``what would have made a difference in this effect?''  A program mutant is a very direct way to answer this question.  The problem with determining causes of events in computer programs is that, as in most problems of causality, too many things can make a difference, most of which are not interesting.  Program mutants are much more likely to be interesting causes for events in a program than changes to test cases, at least for fault identification and debugging purposes, because the practical use of knowing a cause is to change the cause and avoid or produce an effect.  Debugging by changing test cases is not useful; debugging by altering the program is precisely how we seek to use the causes we discover.  Using mutants as causes in order to measure similarity essentially inverts the ideas of David Lewis \cite{LewisCause,LewisCount}, where similarity is used to determine causality.

>>>>

The goal of this paper, an effective metric for computing distance between program executions (in particular, failing test cases), is motivated in large part by progress in automated testing.  Recent advances, some implemented in powerful tools \cite{csmith,jsfunfuzz,ISSTA12,LangFuzz,ZhendongPLDI,ZhendongOOPSLA}, have made it easier to find subtle faults in, e.g., modern optimizing compilers\footnote{In this paper, we take the problem of testing and debugging compilers as representative of the general problem of testing and debugging complex critical systems software.}.  Unfortunately, the blessing of many failing test cases can become a curse for a developer.  Simply discovering which test cases in a large set of failures represent distinct faults (or already known faults) is a very difficult problem \cite{PLDI13,Podgurski04}.  For simple crashes and assertion violations, bucketing failures is sometimes easy.  When a failure is detected by differential testing \cite{Differential}, where two compilers or optimization levels disagree, the problem is so hard that it may turn potential users away from automated test generation \cite{PLDI13}.  Even once a fault is identified, fixing the problem can be very difficult.  Due to complexity of intermediate representation, interaction of optimizations, subtle language semantics, and other issues, debugging compiler faults may often be harder than debugging faults in other kinds of software (this is probably not limited to compilers; debugging systems software in general is hard \cite{mickens}).
We propose the use of program mutants \cite{mutant} to assist debugging based on large sets of automatically generated test cases.  The underlying idea is that if two failures can be repaired in the same way, even if that repair only avoids the failure and does not truly fix it, they are likely due to the same fault. Repairs can also provide  information for understanding and debugging discovered faults. The approach suggests a plausible metric for comparing any two program executions, not just failing test cases.  Comparing executions has been a prominent concept in dynamic analysis throughout much of the last two decades \cite{NearNeighbor,Sumner2011} (e.g., Ball's early discussion of profile similarity \cite{BallConcept}).

@@

\subsection{Comparison by Response to Mutation}

@@

The idea of measuring the distance between failures (or other test cases) by their response to mutants is appealing.  The most important aspect of most applications of distances between executions is some aspect of causality \cite{LewisCause,ZellerBook}.  Since at least the work of Hume \cite{Hume1}, causality has been understood as answering some question of the form ``what would have made a difference in this effect?''  A program mutant is a very direct way to answer this question.  The problem with determining causes of events in computer programs is that, as in most problems of causality, too many things can make a difference, most of which are not interesting.  Program mutants are much more likely to be interesting causes for events in a program than changes to test cases, at least for fault identification and debugging purposes, because the practical use of knowing a cause is to change the cause and avoid or produce an effect.  Debugging by changing test cases is not useful; debugging by altering the program is precisely how we seek to use the causes we discover.  Using mutants as causes in order to measure similarity ``inverts'' the ideas of Lewis \cite{LewisCause,LewisCount}, where similarity determines causality.

### 2016-03-01 metrics.tex
fixes
ready submit

For every failing test case, its response to mutants defines a bit-vector, with
length equal to the number of mutants that repaired any test case,
where a 1 bit indicates that the mutant in question repairs this test
case.  Hamming distance between vectors is not suitable for our purposes.  Consider two test cases, $t_1$ and $t_2$, both repaired
by 100 mutants.  If these two test cases share 90 common repairing
mutants, and disagree for 10, they have a Hamming
distance of 20.  This makes them ``less similar'' than two test cases,
$t_1'$ and $t_2'$, where $t_1'$ is repaired by 5 mutants, $t_2'$ is
repaired by 5 mutants, and \emph{none of those mutants overlap.}   Hamming distance for the first pair of tests is twice as
large, despite the fact we have no evidence $t_1'$ and
$t_2'$ are due to the same fault, in contrast to strong evidence for $t_1$ and $t_2$.  A Jaccard distance, used in some fault
localization algorithms \cite{Liu06}, is a better fit.  For bitvectors $u$ and $v$ of length $n$:
%\vspace{-0.08in}
\[d(u,v) = 
\begin{cases}
\frac{\sum_{i=0}^{n-1} 1\ \text{if}\ u_i \neq v_i\ \text{else}\ 0}{\sum_{i=0}^{n-1} 1
\  \text{if}\ u_i \neq 0 \vee v_i \neq 0\ \text{else}\ 0} & \exists i:u_i\neq 0 \vee v_i\neq 0\\
0 & otherwise
\end{cases}
\]

>>>>

For every failing test case, its response to mutants defines a bit-vector, with
length equal to the number of mutants that repaired any test case,
where a 1 bit indicates that the mutant in question repairs this test
case.  These cannot be effectively compared with Hamming distance. Consider $t_1$ and $t_2$, both repaired
by 100 mutants.  If these two test cases share 90 common repairing
mutants, and disagree for 10, they have a Hamming
distance of 20.  This makes them ``less similar'' than two test cases,
$t_1'$ and $t_2'$, where $t_1'$ is repaired by 5 mutants, $t_2'$ is
repaired by 5 mutants, and \emph{none of those mutants overlap.}   Hamming distance for the first pair of tests is twice as
large, despite the fact we have no evidence $t_1'$ and
$t_2'$ are due to the same fault, in contrast to strong evidence for
$t_1$ and $t_2$.  A Jaccard distance, also used (as similarity) in some fault
localization algorithms \cite{Pinpoint}, is a better fit.  For bitvectors $u$ and $v$ of length $n$:
%\vspace{-0.08in}
\[d(u,v) = 
\begin{cases}
\frac{\sum_{i=0}^{n-1} 1\ \text{if}\ u_i \neq v_i\ \text{else}\ 0}{\sum_{i=0}^{n-1} 1
\  \text{if}\ u_i \neq 0 \vee v_i \neq 0\ \text{else}\ 0} & \exists i:u_i\neq 0 \vee v_i\neq 0\\
0 & otherwise
\end{cases}
\]

### 2016-03-01 relatedwork.tex
ready submit

The most closely related work is that of Chen et. al \cite{PLDI13} (which shares one author with the present paper).  We use two of the same data sets, apply the FPF \cite{Gonzalez} algorithm, and have the same empirical focus of determining how to identify faults for complex compilers.  The primary difference is that the distance metrics used with FPF in the previous work were \emph{ad hoc}, chosen for each program by hand, produced poorer results for the critical portion of the curve for the two hard data sets, and were not extended to provide any fault localization.  This paper expands on Chen et al. by putting the FPF approach in the context of a much richer causal comparison of program executions that applies in theory to any executions, not just failing test cases.  Prior to the FPF-based work, Francis et al. \cite{Podgurski04}, Podgurski et al. \cite{Podgurski03} and others \cite{Liu06,Liblit05} used clustering to attempt to solve similar problems in identifying distinct faults, on simpler human-generated data sets.  The work of Chen et. al \cite{PLDI13} provides an in-depth comparison of clustering and ranking, and the advantages of ranking (as well as some possible limitations).  Jones et al. discuss general issues in debugging multiple faults efficiently \cite{Jones07}.

@@

The use of distance metrics in software engineering includes some of the localization and fault identification efforts discussed above.  Vangala et al. proposed using distance to cluster test cases to improve diversity and eliminate duplicates \cite{VangalaDist}.  Adaptive random testing uses distances (usually between inputs) to choose tests \cite{Chen,ARTChen,ISSTAART}, and Artzi et al. guided test generation for fault localization using path constraint metrics \cite{ArtziDirected}.  Sumner et al. evaluated approaches to generating execution peers for debugging and understanding \cite{Sumner2011} using tree edit distances (and evaluated other distances).  Xin et al. \cite{Xin2008} examined methods for indexing program executions, useful in all methods involving execution alignments.  To our knowledge, most metrics used are essentially spectrum-based \cite{RepsSpectra} (using counts of coverage of code entities), while our metric is essentially causal, based on counterfactual \cite{LewisCause,LewisCount} programs.

>>>>

The most closely related work is that of Chen et. al \cite{PLDI13} (which shares one author with the present paper).  We use two of the same compiler data sets, and both apply the FPF \cite{Gonzalez} algorithm.  The distance metrics used with FPF in the previous work, however, were \emph{ad hoc}, chosen for each program by hand, produced poorer results for the critical portion of the curve for the two hard data sets, and were not extended to provide any fault localization.  This paper uses the FPF approach in the context of a more principled causal comparison of program executions that applies in theory to any executions, not just failing test cases, and introduces novel fault localizations.  Prior to the FPF-based work, Francis et al. \cite{Podgurski04}, Podgurski et al. \cite{Podgurski03} and others \cite{Liu06,Liblit05} used clustering to attempt to solve similar problems in identifying distinct faults, on simpler human-generated data sets.  The work of Chen et. al \cite{PLDI13} provides an in-depth comparison of clustering and ranking, and the advantages of ranking (as well as some possible limitations).  Jones et al. discuss general issues in debugging multiple faults efficiently \cite{Jones07}.

@@

The use of distance metrics in software engineering includes some of the localization and fault identification efforts discussed above (e.g., \cite{NearNeighbor,GroceError,ChakiLev,Liu06}).  Vangala et al. proposed using distance to cluster test cases to improve diversity and eliminate duplicates \cite{VangalaDist}.  Adaptive random testing uses distances (usually between inputs) to choose tests \cite{Chen,ARTChen,ISSTAART}, and Artzi et al. guided test generation for fault localization using path constraint metrics \cite{ArtziDirected}.  Sumner et al. evaluated approaches to generating execution peers for debugging and understanding \cite{Sumner2011} using tree edit distances (and evaluated other distances).  Xin et al. \cite{Xin2008} examined methods for indexing program executions, useful in all methods involving execution alignments.  To our knowledge, most metrics used are essentially spectrum-based \cite{RepsSpectra,BallConcept} (using counts over structural entities), while our metric is essentially causal, based on counterfactual \cite{LewisCause,LewisCount} programs.
